\documentclass[11pt]{amsart}
\usepackage[margin=1.1in]{geometry} 
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}

\usepackage{fancyhdr}
 
\pagestyle{fancy}
\fancyhf{}
\lhead{UC Santa Cruz}
\rhead{Jared G. Galloway}
\rfoot{Page \thepage}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{UC Santa Cruz Biomolecular Engineering and Bioinformatics track Ph.D. Personal History Statement}
\author{Jared G. Galloway}
%\date{}                                           % Activate to display a given date or no date
\newlength{\spacebox}
\settowidth{\spacebox}{8888888888}			% Box to align text
\newcommand{\sepspace}{\vspace*{1em}}		% Vertical space macro

\begin{document}
\maketitle
%\section{}
%\subsection{}

%UC Santa Cruz is interested in a diverse and inclusive graduate student population. 
%In an essay, discuss how your personal background informs your decision to pursue a graduate degree. 
%Include any educational, familial, cultural, economic, or social experiences, challenges, or opportunities relevant to your academic journey; 
%how you might contribute to social or cultural diversity within your chosen field; 
%and/or how you might serve educationally underrepresented segments of society with your degree.

Up until I was offered an undergraduate research position I had worked labor jobs including pest control, painting houses, and moving furniture.
Before that, my father put me to work moving hay, mucking the horse barn, putting up snow fence, and many other strenuous chores.
I strongly contribute my discipline, respect for the trades, and longing to grow intellectually to these tasks.
Being raised on a small ranch in Montana,
I was surrounded by gorgeous landscape and seemingly unlimited access to the outdoors. 
Unfortunately, I was not exposed to a large variety of social, cultural, or political diversity. 
Priding myself on keeping an open mind, I pursued my degree out-of-state in an attempt to gain new perspectives.
This decision has helped me mature in setting which supports diversity, as well as opened my eyes to acknowledging my privilege as a white male in the United States.
From a personal standpoint, I'm pursuing my Ph.D. to 
challenge myself at the academic level, 
and to utilize my position of influence to be an advocate and ally to those who haven't been given the chance to voice their opinion.
\sepspace

While I have never struggled academically I do not contribute my achievements to sheer intelligence. 
Rather, I have been motivated by my attention to detail and strong work ethic.
%THIS NEXT SENTENCE COULD USE SOME WORK
I believe this is among one of the most important attributes as opposed to intellectual ability, which is not fixed from birth but ever expanding as we push ourselves to learn new abstractions for real life applications.
For example, during my undergraduate I came to find a positive correlation between my ability in different flavors of math as I became more familiar with data structures and algorithms. 
Similarly, I believe that my education in information science has laid the foundation for my path towards becoming a bimolecular engineer as well as a bioinformatician.
For this reason I am pursuing my graduate training outside of computer and information science
and will work hard to be interdisciplinary so that I may transport skills which are often applicable across fields. 
\sepspace

While the values instilled in a rural area shaped many of the characteristics I carry today, I believe living in a place of little diversity put me a disadvantage. 
Many of the thought processes I noticed from people in the area were close minded and stubborn.
This type of thinking is not only dangerous, but inhibits learning from, and teaching others with a different viewpoint. 
It is crucial that we give everyone equal opportunity to be heard so that we may benefit from all the advantages diversity brings as well as community growth rather than building barriers.
Ultimately, I believe my privilege carries with it a responsibility to learn from and help allow humans from different cultures and demographics to project their voice.
As a Ph.D. student and a scientist I will use my influence to promote social and cultural diversity by actively pursuing and presenting opportunities for the underrepresented demographics.

\end{document}  